\section{Comparison with \pathcasekegg Web Interface}
\label{sect:kegg_comparison}

\subsection{Features}
\label{sect:kegg_comparison_features}

Table \ref{fig:kegg_comparison_table} shows a comparison of relevant features
from the online and iPad versions of the \pathcasekegg analysis tools. The
similarities and differences are elaborated here.

\begin{table}[ht!]
\centering
\begin{tabular}{ | p{2.75in} | p{1.25in} | p{1.25in} | }
    \hline
                        & Online    & iPad \\ \hline

    Molecule and reaction visualization
                        & Yes       & Yes \\ \hline

    Navigation method   & Four mouse-based navigation tools
                                    & Pan with one finger, zoom with two
                                    \\ \hline

    Activating a subset of reactions based on the organisms a reaction is
    present in
                        & Yes       & Yes \\ \hline

    Moving nodes to rearrange the layout
                        & Yes       & Yes \\ \hline

    Dynamic layout when no frozen layout is present
                        & Yes       & Yes \\ \hline

    Saving layouts for future recall
                        & Yes       & No \\ \hline

    Lists of processes, connected pathways, pathway annotations, and
    pathway-related queries in tabular form
                        & Yes       & No \\ \hline

    Selecting a reaction based on its place in the genome
                        & Yes       & No \\ \hline

    Hiding common molecules
                        & Yes       & No \\ \hline

    Display of data from ENZYME database
                        & No        & Yes \\ \hline

    Display of data from KEGG web site
                        & No        & Yes \\ \hline
\end{tabular}
    \caption{Functional differences between the online \pathcasekegg visualizer
    and the iPad KEGG visualizer}
    \label{fig:kegg_comparison_table}
\end{table}

Functional differences between \keggapp and the online \pathcasekegg tools exist
for reasons similar to the differences between \mawapp and the \pathcasemaw
online tools. In addition to the same changes made by \mawapp, \keggapp allows
the user to rearrange nodes in the visualization, has the ability to generate
dynamic layouts, and adds additional pertinent information from new data
sources.

Some functionality is missing due to development time limitations. Selection of
reactions by where they are located in the genome was not implemented for this
reason, as well as the ability to save a manually arranged layout for future
recall.

Other functionality is missing because web services have not been implemented
for them. Database queryies are missing for this reason, though web services to
support them are being developed.
