\section{Literature Survey}

The literature I have found falls into three major categories. The first category is general multitouch interface research. Not much of the general multitouch research seems to apply, since iOS already has a standard interface and guidelines for implementing new interfaces. All custom interactions happen within the graph view. However, some relevant research is presented here for completeness.

A few papers present taxonomies of gestures. Frisch et al provide one example \cite{multitouch:gestures}. (I am looking for more.)

Others offer new interfaces to help bridge the divide between multitouch input devices and the pre-iPad desktop interfaces. In this vein, Benko et al describe a new interface to help a user select very small pointing targets \cite{multitouch:tiny-select}. However, modern touch interfaces circumvent this issue by not requiring users to select unreasonably tiny targets in the first place.

The second category is graph interface research. Kobourov et al have a multi-touch interface for making pairs of graphs with similarly placed nodes \cite{graph-interaction:simultaneousgraphdrawing}, which will likely be useful in the SMDA application. I have found one paper in which the researchers use bubbles along the edges of the screen to represent offscreen nodes (but I am so far unable to download it).

There is a lot of talk about UML, but it is not relevant to this application. Another paper uses a hybrid graphical and text interface to explore graphs, but text interfaces do not work well on touch-based devices. There may be more relevant papers in this area, but I am still sifting through them.

The third category is graph theory. In particular, I am interested in graph rewriting, i.e. turning one graph into another based on rules. I found a good starting paper, but the version I have is corrupted. There are also many papers on graph query languages.

I am still looking for papers at the intersection between graph query languages, graph rewriting, and multitouch interfaces. For example, how can a user express a similarity query without entering any text? Or, how to easily select graphs that match a pattern?
