The MAW app is split into two major components. The first is a pathway browser
that adds tissue and compartment information to the individual pathways. The
second is a touch-based implementation of the SMDA (Steady-State Metabolic
Network Dynamics Analysis) tool.

\section{Pathway Browser}

The MAW iPad application displays metabolic pathway graphs for all pathways for
which the PathCase MAW database provides a static layout. These graphs can be
navigated by panning and zooming.

\input{maw/interface}

\input{maw/implementation}

\section{SMDA}
\label{sect:smda}

SMDA (Steady-State Metabolic Network Dynamics Analysis) is a tool to analyze
metabolomics data in terms of the dynamic behavior of the metabolic network.
Given a set of metabolite measurements, it identifies the metabolic mechanisms
that lead to changes in the concentrations of given metabolites, and interprets
the metabolic consequences of the observed changes in terms of physiological
problems, nutritional deficiencies, or diseases.

The input data consists of one or more pathways, zero or more transport
processes, zero or more reactions, and measurements of the amounts of specific
metabolites within the pathways. This input forms a \textbf{sub-network}.

The output is a set of all possible scenarios of the sub-network based on the
input measurements. In a scenario, each reaction is labeled as either active or
inactive.

The iPad version of this tool allows the user to complete both the input and
the output step. Input is done through a table-based user interface. The results
are shown as a series of graph views that can be paged through.

\input{maw/smda_interface}

\input{maw/smda_implementation}

\input{maw/web_comparison}

\input{maw/future}
